---
title: "full stack data-science finance (small) project"
author: "Olivier Bauthéac"
output: github_document
---



# preprocessing (ELT)

## extract

### minimum required

In an excel woorkbook, query Bloomberg for historical (bdh) as well as contemporaneous (bdp) data for a market index as well as a broad cross-section of U.S. stocks. Historical data should be retrieved from October 1^st^ 2016 to today at the daily frequency on individual ticker specific sheets (one sheet per name). All names' contemporaneous data, on the other hand, should sit on a single sheet. The Bloomberg ticker for the market index is 'RAY Index' while those for the corporation names are listed below:

BBG stock tickers |                 |                 |                 |                 
------------------|-----------------|-----------------|-----------------|----------------
ADM US Equity     | CIVI US Equity  | GBX US Equity   | LIND US Equity  | SERV US Equity
AE US Equity      | CLGX US Equity  | GDI US Equity   | LZB US Equity   | SGA US Equity
AGCO US Equity    | CLR US Equity   | GHC US Equity   | MAN US Equity   | SITE US Equity
AJRD US Equity    | COMM US Equity  | GME US Equity   | MEI US Equity   | SMP US Equity
ALG US Equity     | CRL US Equity   | GOLF US Equity  | MLR US Equity   | SPXC US Equity
AMD US Equity     | CTB US Equity   | GPN US Equity   | MRC US Equity   | STRT US Equity
AMOT US Equity    | CTLT US Equity  | GTLS US Equity  | MTD US Equity   | SUPN US Equity
ASGN US Equity    | CTXS US Equity  | HFC US Equity   | MTZ US Equity   | TAST US Equity
ATRO US Equity    | DHI US Equity   | HOFT US Equity  | NC US Equity    | TMO US Equity
AVT US Equity     | DKS US Equity   | HPE US Equity   | NGVT US Equity  | TNET US Equity
AWI US Equity     | EBIX US Equity  | HURC US Equity  | NHC US Equity   | TPB US Equity
BBBY US Equity    | EEFT US Equity  | HWKN US Equity  | NUE US Equity   | UBNT US Equity
BFAM US Equity    | ELF US Equity   | HY US Equity    | OSIS US Equity  | UFPI US Equity
BID US Equity     | ELVT US Equity  | IAC US Equity   | OSK US Equity   | UFS US Equity
BIG US Equity     | EML US Equity   | IART US Equity  | PFGC US Equity  | USAK US Equity
BKNG US Equity    | ENTG US Equity  | IBP US Equity   | PGTI US Equity  | VLGEA US Equity
BLD US Equity     | ERI US Equity   | IDTI US Equity  | PKI US Equity   | VLO US Equity
BSET US Equity    | ETH US Equity   | INT US Equity   | PLPC US Equity  | VRSK US Equity
BWA US Equity     | FICO US Equity  | IOSP US Equity  | PRAH US Equity  | WBC US Equity
BYD US Equity     | FISV US Equity  | ITRI US Equity  | PSX US Equity   | WERN US Equity
CAL US Equity     | FL US Equity    | JLL US Equity   | RBC US Equity   | WGO US Equity
CBRE US Equity    | FLR US Equity   | KHC US Equity   | RS US Equity    | WRK US Equity
CENTA US Equity   | FLT US Equity   | KSU US Equity   | RXN US Equity   | XPO US Equity
CHEF US Equity    | FTV US Equity   | LGND US Equity  | SCL US Equity   | ZBRA US Equity


The historical time series should include the following market & book data fields:

field                 | Bloomberg symbol                       
----------------------|-----------------
close price           | PX_LAST
book value per share  | BOOK_VAL_PER_SH
earnings per share    | TRAIL_12M_EPS
dividend per share    | TRAIL_12M_DVD_PER_SH
debt                  | SHORT_AND_LONG_TERM_DEBT
equity                | TOTAL_EQUITY
current assets        | BS_CUR_ASSET_REPORT
current liabilities   | BS_CUR_LIAB
sales                 | SALES_REV_TURN

Contemporaneous data on the other hand should include the number of shares outstanding, number of directors on the board, number of women on the board, number of board meetings per year, long company name and companie description. Explore Bloomberg to find the corresponding field symbols.


### going further

* Using VBA, make your workbook updatable. Ammend your workbook so that it retrieves up to date data in one clic.
  + Hint 1. Update doesn't necessarily mean adding most recent values to an existing time series. Requerying the whole data up to the most recent date would work as well.
  + Hint 2. Inspect the BQL syntax in Bloomberg formula cells, ammend accordingly.

* Using VBA, make your workbook flexible. Ammend your workbook so that it can retrieve data for any set of stocks/indexes & market/book fields at various frequencies (year, month, week, day), from and to any date.
  + Hint 1. Object oriented programming could help; excel table objects in particular.
  + Hint 2. Create an update sheet with tickers list, parameters (frequency, start and end dates) and fields. This sheet could also be use to host the contemporaneous dataset.

* Using VBA, make your workbook fully portable. If you open your workbook without a live Bloomberg connection you'll notice you loose the contemporaneous dataset; try to fix that problem somehow. 
  + Hint 1. VBA events could help.
  

You now have a fully customizable & portable tool to retrieve financial data from Bloomberg and now it's time to use it.



## Load

```{r, include = FALSE}
library(readxl); library(dplyr); library(data.table)

path <- 'data-raw/pullit.xlsm'; ticker_market <- 'RAY Index'; window <- 252L

ticker_firms <- read_excel(path, sheet = "update", range = cell_cols("D")) %>% filter(ticker != ticker_market)

static <- read_excel(path, sheet = "update", range = cell_cols("D:H")) %>% 
  filter(ticker != ticker_market) %>% mutate_at(vars(-ticker), funs(as.numeric))

tickers <- gsub(pattern = '/', replacement = '~', x = c(ticker_market, ticker_firms$ticker))

historical <- lapply(tickers, function(x) read_excel(path, sheet = x, skip = 1L) %>% 
                       mutate(Date = as.Date(Date), ticker = x) %>% 
                       mutate_at(vars(-c('ticker', 'Date')), funs(as.numeric)) %>% 
                       filter(! is.na(Date))) %>% 
  rbindlist()
```

In R or Python (examplesolutions will be provided for both programming languages), load the workbook data in memory. Organise the data in two dataframes, one for the historical times series, the other for static (contemporaneous) data. The time series dataframe should have a two-level row index including tickers & dates while columns should host the corresponding time series; the dataframe structure should look as follow:

```{r, message = FALSE, echo = FALSE}
as_tibble(select(historical, ticker, everything()))
```

The static dataset on the other hand should be row-indexed by tickers and have columns hosting the corresponding static fields; the dataframe structure should look as follow:

```{r, message = FALSE, echo = FALSE}
as_tibble(static)
```


## Transform







